{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2023-11-01 paper/si.tex
swaps the notation for the imitation probability and fixation probability

The long time scale describes the evolutionary dynamics. 
Here, players are allowed to update their strategies based on the payoffs they yield. 
We model these strategy updates with a pairwise comparison process~\citep{traulsen2007pairwise}.
This process assumes that at regular time intervals, one player is randomly selected from the population.
We refer to this player as the `learner' (\learner). 
The learner is then given an opportunity to update its strategy. 
There are two possibilities for how this update may occur. 
With probability $\mu$, the player's strategy mutates randomly. 
In that case, the player's new strategy is drawn uniformly from the space of all reactive strategies~$[0,1]^3$.  
With probability $1\!-\!\mu$, the player compares itself with a neighbor. 
In that case, the player randomly picks another individual from the population (referred to as the `role model', \rolemodel). 
The learner adopts the role model's strategy with a probability \(\rho\), given by
\begin{equation} \label{Eq:rho}
    \rho(\pi_{\learner}, \pi_{\rolemodel}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
\end{equation}
The parameter \(\beta\) is the selection strength.
It determines how important payoff differences are for the learner's decision to imitate the role model. 
The variables $\pi_\rolemodel$ and $\pi_\learner$ refer to the relevant payoffs of the role model and the learner, respectively. 
The exact value of these payoffs depend on the players' memory. 
We say players have {\it perfect memory} when $\pi_\rolemodel$ and $\pi_\learner$ are given by the players' expected payoffs (across all rounds and across all possible co-players). 
We say players have {\it limited memory} when $\pi_\rolemodel$ and $\pi_\learner$ are given by the players' realized payoff in the very last round of the game with their very last interaction partner. 
In addition, we consider several model extensions in which individuals have memory capacities in between these two extremes. 
We provide a detailed description of these different settings and the resulting payoffs in Section~\ref{section:analyticalresults}.\\

@@

Initially, the process starts with a population where all members use the same strategy (referred to as the resident strategy, \resident). 
Then one individual adopts a mutant strategy selected uniformly at random from the set of feasible strategies.
The fixation probability \(\varphi_{\mutant}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
\begin{equation}\label{eq:fixation_probability}
    \varphi_{\mutant} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_{k=1}^i \frac{\lambda^-_k}{\lambda^+_k}}.
\end{equation}
Here, the index $k$ corresponds to the current number of players with the mutant strategy (mutants). 
The variables \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of mutants decreases or increases within a single updating step. 
These probabilities depend on the probability that a mutant and a resident are chosen as the learner and the resident, respectively. 
In addition, they depend on the respective switching probability~$\rho$, as described by Eq.~\eqref{Eq:rho}. 
We specify the exact values of  \(\lambda^-_k, \lambda^+_k\) for each memory-setting in the next section.

Depending on the fixation probability \(\varphi_{\mutant}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
Afterwards, another random mutant strategy is introduced into the population. 
We iterate this elementary population updating process for a large number of mutant strategies. 
At each step, we record the current resident strategy and the resulting average cooperation rate.

@@

\begin{algorithm}[t]
  \SetAlgoLined
  $N \leftarrow$ population size\;
  %$k \leftarrow 1$\; 
  resident $\leftarrow$ starting resident\;
   \While{ $t <$ maximum number of steps}
   {mutant $\leftarrow$ random strategy\;
   fixation probability $\leftarrow \varphi_\mutant $\;
   \If{$\varphi_{\mutant} >$ random: $i \rightarrow [0,1]$}
   {resident $\leftarrow$ mutant;}}
   \caption{Evolutionary process in the limit of rare mutations}\label{algorithm:pairwise_comparison}
\end{algorithm}

@@

~\\
\noindent
After these preparations, consider now a population with \(k\) mutants and \(N - k\) residents, whose strategies we denote by \(\strategy_\mutant=(y_\mutant,p_\mutant, q_\mutant)\) and \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\), respectively.
Assuming that population members are matched randomly (or equivalently, that they interact with all other population members), the resulting expected payoffs of residents and mutants are
\begin{equation} \label{Eq:ExpPay}
  \begin{array}{lcrcr}
  \displaystyle \pi_\resident(k)& = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \pi(\strategy_\resident,\strategy_\resident)	&+	&\displaystyle\frac{k}{N-1}\cdot \pi(\strategy_\resident,\strategy_\mutant),\\[0.5cm]
  \displaystyle \pi_\mutant(k)& = &\displaystyle\frac{N-k}{N-1}\cdot \pi(\strategy_\mutant,\strategy_\resident) &+	&\displaystyle\frac{k-1}{N-1}\cdot \pi(\strategy_\mutant,\strategy_\mutant).\\
  \end{array}
\end{equation}
The number of mutants in the population decreases in a single time step if a mutant is chosen to be the learner and adopts the strategy
of a resident. 
Similarly, it increases if a resident is the learner and adopts the strategy of a mutant. 
The respective transition probabilities are 
\begin{align*}
  \lambda^-_k \!=\! \frac{N-k}{N}\frac{k}{N-1}\rho\big(\pi_\mutant(k),\,\pi_\resident(k)\big) \quad \text{ and } \quad \lambda^+_k \!=\!\frac{k}{N}\frac{N-k}{N-1}\rho\big(\pi_\resident(k),\,\pi_\mutant(k)\big).
\end{align*}
For $\rho$ as defined by Eq.~\eqref{Eq:rho}, the ratio of these two transition probabilities simplifies to 
\begin{equation} \label{eq:LambdaRatio}
 \frac{\lambda^-_k}{\lambda^+_k} \!=\!  \frac{\rho\big(\,\pi_\mutant(k),~\pi_\resident(k)\,\big)}{\rho\big(\,\pi_\resident(k),~\pi_\mutant(k)\,\big)} 
 =e^{-\beta\big(\pi_\mutant(k)-\pi_\resident(k)\big)}.
\end{equation}
Based on these ratios for each $k$, we can compute the mutant's fixation probability by Eq.~\eqref{eq:fixation_probability}.

@@

\noindent
The first term on the right side corresponds to the case that the learner and
the role model happened to be matched directly for their last interaction. 
In that case, only those payoff pairs can occur that are feasible in a direct interaction. 
That is, it needs to be the case that $(u_\resident,u_\mutant)\in \mathcal{U}_F$, as represented by the respective indicator function. 
The probability that the number of mutants increase or decreases by one is now given by
\begin{equation}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant\,\in\,\mathcal{U}} x(u_\resident,u_\mutant)\cdot \rho(u_\resident,u_\mutant), \\[0.5cm]
\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant \,\in\, \mathcal{U}} x(u_\resident,u_\mutant)\cdot \rho(u_\mutant,u_\resident).
\end{array}
\end{equation}

@@

\noindent
As a result, the transition probabilities for the number of mutants is now given by
\begin{equation}\label{eq:ratio_TwoInteractions}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})\cdot \rho\left(\frac{u_{\resident 1} \!+\! u_{\resident 2}}{2}, \frac{u_{\mutant 1} \!+\! u_{\mutant 2}}{2}\right), \\[0.5cm]
\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2},u_{\mutant 1}u_{\mutant 2})\cdot \rho\left(\frac{u_{\mutant 1} \!+\! u_{\mutant 2}}{2}, \frac{u_{\resident 1} \!+\! u_{\resident 2}}{2}\right).
\end{array}
\end{equation}
Note that we assume players to use their average payoff across their two interactions for comparisons.\\

@@

 ~\\
 \noindent
 {\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
 To compute the ratio of transition probabilities, let us again take a population perspective. 
 We consider $k$ mutants with strategy $\strategy_\mutant=(y_\mutant,p_\mutant,q_\mutant)$ and $N\!-\!k$ residents with strategy $\strategy_\resident = (y_\resident,p_\resident,q_\mutant)$. 
Without loss of generality, we assume that either the learner or the role model is a resident, and that the respective other player is a mutant. 
Let $x(u_\resident u'_\resident, u_\mutant u'_\mutant)$ be the probability that the two players received payoffs $u$ and $u'$ in the last two rounds of their last repeated game.
We can compute this probability based on the same logic as in the limited-memory setting (Section~\ref{section:limited_memory}). 
This yields
\begin{equation*}\label{eq:Chi_TwoRounds} 
\setlength{\arraycolsep}{1pt} 
\begin{array}{llll}
x(u_\resident u'_\resident, u_\mutant u'_\mutant)	 &=
&\frac{1}{N\!-\!1}\cdot  &w_{u_\resident u_\resident'}\!(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_\resident,u_\mutant)\in \mathcal{U}_F}\cdot 1_{(u'_\resident,u'_\mutant)\in \mathcal{U}_F}\\[0.5cm]
&+	
&\frac{N\!-\!2}{N\!-\!1} \cdot 
&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\mutant) \,w_{u_\mutant u_\mutant'} \!(\strategy_\mutant,\strategy_\mutant) + 
 \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\mutant) \,w_{u_\mutant u_\mutant'}  \!(\strategy_\mutant,\strategy_\resident)\right.\\[0.5cm]
&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\resident) \,w_{u_\mutant u_\mutant'} \!(\strategy_\mutant,\strategy_\mutant) + 
 \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\resident) \,w_{u_\mutant u_\mutant'}  \!(\strategy_\mutant,\strategy_\resident)\right].
\end{array}
\end{equation*}
Overall, we obtain the following formula for the transition probabilities 
\begin{equation}\label{eq:ratio_TwoRounds}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \rho\left(\frac{u_\resident \!+\! u_\resident'}{2}, \frac{u_\mutant \!+\! u_\mutant '}{2}\right), \\[0.5cm]
\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \rho\left( \frac{u_\mutant \!+\! u_\mutant '}{2}, \frac{u_\resident \!+\! u_\resident'}{2}\right).
\end{array}
\end{equation}

@@

As our final extension, we consider a combination of the previous two settings, discussed in Sections~\ref{section:m_one_n_two} and \ref{section:m_two_n_one}. 
Now, individuals take into account two interactions with different co-players, and from each interaction, they take into account the last two rounds. 
The probability that the number of mutants increases or decreases by one is now given by
\begin{equation}\label{eq:ratio_TwoRounds}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} 
\hspace{-0.5cm}
\sum_
{\scriptsize \begin{array}{l}
u_{\resident 1},u_{\resident 1}', u_{\resident 2},u_{\resident 2}'\\
u_{\mutant 1},u_{\mutant 1}', u_{\mutant 2},u_{\mutant 2}'
\end{array}} 
\hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
\rho\!\left(\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4},
\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4}\!\right), \\[1.4cm]

\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} 
\hspace{-0.5cm}
\sum_
{\scriptsize \begin{array}{l}
u_{\resident 1},u_{\resident 1}', u_{\resident 2},u_{\resident 2}'\\
u_{\mutant 1},u_{\mutant 1}', u_{\mutant 2},u_{\mutant 2}'
\end{array}} 
\hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
\rho\!\left(\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4},
\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4}\!\right).
\end{array}
\end{equation}
 Though we do not pursue any further analytical exploration of this case, in the
next section we present simulation results for this setting, as well as for the other settings.\\

@@

\begin{equation*}
  \setlength{\arraycolsep}{1pt} 
  \begin{array}{llll}
    \lambda^+	 &=
  &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)\right)\\[0.5cm]
  &+
  &\frac{N\!-\!2}{N\!-\!1} \cdot 
  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \rho(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\mutant)) + 
   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \rho(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)) \right.\\[0.5cm]
  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \rho(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\mutant)) + 
   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \rho(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\resident))\right],
  \end{array}
\end{equation*}

>>>>

The long time scale describes the evolutionary dynamics. 
Here, players are allowed to update their strategies based on the payoffs they yield. 
We model these strategy updates with a pairwise comparison process~\citep{traulsen2007pairwise}.
This process assumes that at regular time intervals, one player is randomly selected from the population.
We refer to this player as the `learner' (\learner). 
The learner is then given an opportunity to update its strategy. 
There are two possibilities for how this update may occur. 
With probability $\mu$, the player's strategy mutates randomly. 
In that case, the player's new strategy is drawn uniformly from the space of all reactive strategies~$[0,1]^3$.  
With probability $1\!-\!\mu$, the player compares itself with a neighbor. 
In that case, the player randomly picks another individual from the population (referred to as the `role model', \rolemodel). 
The learner adopts the role model's strategy with a probability \(\varphi\), given by
\begin{equation} \label{Eq:rho}
    \varphi(\pi_{\learner}, \pi_{\rolemodel}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
\end{equation}
The parameter \(\beta\) is the selection strength.
It determines how important payoff differences are for the learner's decision to imitate the role model. 
The variables $\pi_\rolemodel$ and $\pi_\learner$ refer to the relevant payoffs of the role model and the learner, respectively. 
The exact value of these payoffs depend on the players' memory. 
We say players have {\it perfect memory} when $\pi_\rolemodel$ and $\pi_\learner$ are given by the players' expected payoffs (across all rounds and across all possible co-players). 
We say players have {\it limited memory} when $\pi_\rolemodel$ and $\pi_\learner$ are given by the players' realized payoff in the very last round of the game with their very last interaction partner. 
In addition, we consider several model extensions in which individuals have memory capacities in between these two extremes. 
We provide a detailed description of these different settings and the resulting payoffs in Section~\ref{section:analyticalresults}.\\

@@

Initially, the process starts with a population where all members use the same strategy (referred to as the resident strategy, \resident). 
Then one individual adopts a mutant strategy selected uniformly at random from the set of feasible strategies.
The fixation probability \(\rho_{\mutant}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
\begin{equation}\label{eq:fixation_probability}
    \rho_{\mutant} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_{k=1}^i \frac{\lambda^-_k}{\lambda^+_k}}.
\end{equation}
Here, the index $k$ corresponds to the current number of players with the mutant strategy (mutants). 
The variables \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of mutants decreases or increases within a single updating step. 
These probabilities depend on the probability that a mutant and a resident are chosen as the learner and the resident, respectively. 
In addition, they depend on the respective switching probability~$\varphi$, as described by Eq.~\eqref{Eq:rho}. 
We specify the exact values of  \(\lambda^-_k, \lambda^+_k\) for each memory-setting in the next section.

Depending on the fixation probability \(\rho_{\mutant}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
Afterwards, another random mutant strategy is introduced into the population. 
We iterate this elementary population updating process for a large number of mutant strategies. 
At each step, we record the current resident strategy and the resulting average cooperation rate.

@@

\begin{algorithm}[t]
  \SetAlgoLined
  $N \leftarrow$ population size\;
  %$k \leftarrow 1$\; 
  resident $\leftarrow$ starting resident\;
   \While{ $t <$ maximum number of steps}
   {mutant $\leftarrow$ random strategy\;
   fixation probability $\leftarrow \rho_\mutant $\;
   \If{$\rho_{\mutant} >$ random: $i \rightarrow [0,1]$}
   {resident $\leftarrow$ mutant;}}
   \caption{Evolutionary process in the limit of rare mutations}\label{algorithm:pairwise_comparison}
\end{algorithm}

@@

~\\
\noindent
After these preparations, consider now a population with \(k\) mutants and \(N - k\) residents, whose strategies we denote by \(\strategy_\mutant=(y_\mutant,p_\mutant, q_\mutant)\) and \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\), respectively.
Assuming that population members are matched randomly (or equivalently, that they interact with all other population members), the resulting expected payoffs of residents and mutants are
\begin{equation} \label{Eq:ExpPay}
  \begin{array}{lcrcr}
  \displaystyle \pi_\resident(k)& = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \pi(\strategy_\resident,\strategy_\resident)	&+	&\displaystyle\frac{k}{N-1}\cdot \pi(\strategy_\resident,\strategy_\mutant),\\[0.5cm]
  \displaystyle \pi_\mutant(k)& = &\displaystyle\frac{N-k}{N-1}\cdot \pi(\strategy_\mutant,\strategy_\resident) &+	&\displaystyle\frac{k-1}{N-1}\cdot \pi(\strategy_\mutant,\strategy_\mutant).\\
  \end{array}
\end{equation}
The number of mutants in the population decreases in a single time step if a mutant is chosen to be the learner and adopts the strategy
of a resident. 
Similarly, it increases if a resident is the learner and adopts the strategy of a mutant. 
The respective transition probabilities are 
\begin{align*}
  \lambda^-_k \!=\! \frac{N-k}{N}\frac{k}{N-1}\varphi\big(\pi_\mutant(k),\,\pi_\resident(k)\big) \quad \text{ and } \quad \lambda^+_k \!=\!\frac{k}{N}\frac{N-k}{N-1}\varphi\big(\pi_\resident(k),\,\pi_\mutant(k)\big).
\end{align*}
For $\varphi$ as defined by Eq.~\eqref{Eq:rho}, the ratio of these two transition probabilities simplifies to 
\begin{equation} \label{eq:LambdaRatio}
 \frac{\lambda^-_k}{\lambda^+_k} \!=\!  \frac{\varphi\big(\,\pi_\mutant(k),~\pi_\resident(k)\,\big)}{\varphi\big(\,\pi_\resident(k),~\pi_\mutant(k)\,\big)} 
 =e^{-\beta\big(\pi_\mutant(k)-\pi_\resident(k)\big)}.
\end{equation}
Based on these ratios for each $k$, we can compute the mutant's fixation probability by Eq.~\eqref{eq:fixation_probability}.

@@

\noindent
The first term on the right side corresponds to the case that the learner and
the role model happened to be matched directly for their last interaction. 
In that case, only those payoff pairs can occur that are feasible in a direct interaction. 
That is, it needs to be the case that $(u_\resident,u_\mutant)\in \mathcal{U}_F$, as represented by the respective indicator function. 
The probability that the number of mutants increase or decreases by one is now given by
\begin{equation}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant\,\in\,\mathcal{U}} x(u_\resident,u_\mutant)\cdot \varphi(u_\resident,u_\mutant), \\[0.5cm]
\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant \,\in\, \mathcal{U}} x(u_\resident,u_\mutant)\cdot \varphi(u_\mutant,u_\resident).
\end{array}
\end{equation}

@@

\noindent
As a result, the transition probabilities for the number of mutants is now given by
\begin{equation}\label{eq:ratio_TwoInteractions}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})\cdot \varphi\left(\frac{u_{\resident 1} \!+\! u_{\resident 2}}{2}, \frac{u_{\mutant 1} \!+\! u_{\mutant 2}}{2}\right), \\[0.5cm]
\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2},u_{\mutant 1}u_{\mutant 2})\cdot \varphi\left(\frac{u_{\mutant 1} \!+\! u_{\mutant 2}}{2}, \frac{u_{\resident 1} \!+\! u_{\resident 2}}{2}\right).
\end{array}
\end{equation}
Note that we assume players to use their average payoff across their two interactions for comparisons.\\

@@

 ~\\
 \noindent
 {\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
 To compute the ratio of transition probabilities, let us again take a population perspective. 
 We consider $k$ mutants with strategy $\strategy_\mutant=(y_\mutant,p_\mutant,q_\mutant)$ and $N\!-\!k$ residents with strategy $\strategy_\resident = (y_\resident,p_\resident,q_\mutant)$. 
Without loss of generality, we assume that either the learner or the role model is a resident, and that the respective other player is a mutant. 
Let $x(u_\resident u'_\resident, u_\mutant u'_\mutant)$ be the probability that the two players received payoffs $u$ and $u'$ in the last two rounds of their last repeated game.
We can compute this probability based on the same logic as in the limited-memory setting (Section~\ref{section:limited_memory}). 
This yields
\begin{equation*}\label{eq:Chi_TwoRounds} 
\setlength{\arraycolsep}{1pt} 
\begin{array}{llll}
x(u_\resident u'_\resident, u_\mutant u'_\mutant)	 &=
&\frac{1}{N\!-\!1}\cdot  &w_{u_\resident u_\resident'}\!(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_\resident,u_\mutant)\in \mathcal{U}_F}\cdot 1_{(u'_\resident,u'_\mutant)\in \mathcal{U}_F}\\[0.5cm]
&+	
&\frac{N\!-\!2}{N\!-\!1} \cdot 
&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\mutant) \,w_{u_\mutant u_\mutant'} \!(\strategy_\mutant,\strategy_\mutant) + 
 \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\mutant) \,w_{u_\mutant u_\mutant'}  \!(\strategy_\mutant,\strategy_\resident)\right.\\[0.5cm]
&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\resident) \,w_{u_\mutant u_\mutant'} \!(\strategy_\mutant,\strategy_\mutant) + 
 \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\resident) \,w_{u_\mutant u_\mutant'}  \!(\strategy_\mutant,\strategy_\resident)\right].
\end{array}
\end{equation*}
Overall, we obtain the following formula for the transition probabilities 
\begin{equation}\label{eq:ratio_TwoRounds}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \varphi\left(\frac{u_\resident \!+\! u_\resident'}{2}, \frac{u_\mutant \!+\! u_\mutant '}{2}\right), \\[0.5cm]
\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \varphi\left( \frac{u_\mutant \!+\! u_\mutant '}{2}, \frac{u_\resident \!+\! u_\resident'}{2}\right).
\end{array}
\end{equation}

@@

As our final extension, we consider a combination of the previous two settings, discussed in Sections~\ref{section:m_one_n_two} and \ref{section:m_two_n_one}. 
Now, individuals take into account two interactions with different co-players, and from each interaction, they take into account the last two rounds. 
The probability that the number of mutants increases or decreases by one is now given by
\begin{equation}\label{eq:ratio_TwoRounds}
\begin{array}{l}
\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} 
\hspace{-0.5cm}
\sum_
{\scriptsize \begin{array}{l}
u_{\resident 1},u_{\resident 1}', u_{\resident 2},u_{\resident 2}'\\
u_{\mutant 1},u_{\mutant 1}', u_{\mutant 2},u_{\mutant 2}'
\end{array}} 
\hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
\varphi\!\left(\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4},
\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4}\!\right), \\[1.4cm]

\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} 
\hspace{-0.5cm}
\sum_
{\scriptsize \begin{array}{l}
u_{\resident 1},u_{\resident 1}', u_{\resident 2},u_{\resident 2}'\\
u_{\mutant 1},u_{\mutant 1}', u_{\mutant 2},u_{\mutant 2}'
\end{array}} 
\hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
\varphi\!\left(\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4},
\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4}\!\right).
\end{array}
\end{equation}
 Though we do not pursue any further analytical exploration of this case, in the
next section we present simulation results for this setting, as well as for the other settings.\\

@@

\begin{equation*}
  \setlength{\arraycolsep}{1pt} 
  \begin{array}{llll}
    \lambda^+	 &=
  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)\right)\\[0.5cm]
  &+
  &\frac{N\!-\!2}{N\!-\!1} \cdot 
  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \varphi(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\mutant)) + 
   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)) \right.\\[0.5cm]
  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \varphi(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\mutant)) + 
   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \varphi(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\resident))\right],
  \end{array}
\end{equation*}

\begin{equation*}
  \setlength{\arraycolsep}{1pt} 
  \begin{array}{llll}
    \lambda^-	 &=
  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
  &+
  &\frac{N\!-\!2}{N\!-\!1} \cdot 
  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \varphi(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\mutant)) + 
   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.5cm]
  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \varphi(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\resident)) + 
   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \varphi(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\resident),)\right].
  \end{array}
\end{equation*}
